---
title: "Practical3Exercise"
author: "2140365 Yanlin Yu"
date: "2023-10-20"
output: html_document
---

```{r LoadLib, echo=FALSE}
library(dplyr)
library(rio)
library(magrittr)
library(readxl)
StopS <- read_excel("StopSearchWk3.xlsx")
names(StopS)   # check what is in the file.
```


```{r GetCols,echo=FALSE}
StopS <- StopS[,c(1,2,7:15)]
names(StopS)
```


```{r RenameCols, echo=FALSE}
names(StopS)[c(4:6,8,10,11)] <- c('AgeRange', 'SelfEthnicity', 'OfficerDefEthnic', 'AimSearch','OutcomeLinkAim','RemovalMoreClothing')
names(StopS)
```


```{r ShowTableDetail, echo=FALSE}
str(StopS)
```


```{r}
StopS <- StopS %>% mutate(
  Gender=factor(Gender), Type=factor(Type), AgeRange=factor(AgeRange),
  SelfEthnicity=factor(SelfEthnicity),OfficerDefEthnic=factor(OfficerDefEthnic),
  Legislation=factor(Legislation),AimSearch=factor(AimSearch),Outcome=factor(Outcome),
)
str(StopS)
```


```{r FactorSummary, echo=FALSE}
summary(StopS)
```


```{r TypeTable, echo=FALSE}
 Types <- factor(StopS$Type,labels=c("PersonVehicle","Person"))
 table(Types,StopS$Type) # This is an example of checking you have got it right!
```


```{r}
StopS <- StopS %>% mutate(Type=factor(StopS$Type,labels=c("PersonVehicle","Person")) )
summary(StopS)
```


```{r SetLabels, echo=FALSE}
StopS <- StopS %>% mutate(Legislation=factor(as.numeric(Legislation), 
                                              levels=c(1,2,3,4),                                             labels=c("JusticeFirearms","JusticeFirearms","DrugsMisuse","Evidence")))
summary(StopS)
```


```{r Lubridate, echo=FALSE}
library(lubridate)
 StopS$dayDate = date(StopS$Date)
StopS$hourDay = as.numeric(substring(StopS$Date,12,13)) # Find a better way?
save(StopS, file="StopSearch.rda") # save for future use.
```


```{r TableFunc, echo=FALSE}
table(StopS$Gender,StopS$AgeRange)
 table(StopS$AgeRange,StopS$Gender)
 proportions(table(StopS$Gender,StopS$AgeRange),1)
 100*proportions(table(StopS$Gender,StopS$AgeRange),2)
```


```{r KableFunc, echo=FALSE}
knitr::kable(table(StopS$Gender,StopS$AgeRange),
  caption = "Gender by Age; 401 people stopped")
knitr::kable(100*proportions(table(StopS$Gender,StopS$AgeRange),2),
             digits=0,
  caption = "Percentage sex by age; 401 people stopped")
knitr::kable(100*proportions(table(StopS$Gender,StopS$AgeRange),2),
             digits=1,
  caption = "Percentage sex by age; 401 people stopped")
knitr::kable(100*proportions(table(StopS$Gender,StopS$AgeRange),1),
             digits=1,
  caption = "Percentage sex by age; 401 people stopped")
```


```{r PlotAgeDistn, echo=FALSE}
 plot(table(StopS$Legislation,StopS$AgeRange),
      main="Distribution of age by legislation for 401 people stopped")
```


```{r PlotAnother, echo=FALSE}
 plot(table(StopS$Legislation,StopS$AgeRange,StopS$Gender),
      main="Distribution of age by legislation for 401 people stopped")
```


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
```


```{r, echo=FALSE}
 svg(file = "JLH.svg", width = 8, height = 5)
 print( plot(table(StopS$Legislation,StopS$AgeRange),
      main="Distribution of age by legislation for 401 people stopped"))
```


```{r, echo=FALSE}
 ##  This writes the graph to the file
 invisible(dev.off())          ##  This is essential, to close the file
```


```{r importGGplot, echo=FALSE}
library(ggplot2)
```


```{r}
age_Legislation <- StopS %>%
  group_by(AgeRange, Legislation)
age_Legislation
```



```{r age_legislation_distn_figure, echo=FALSE}
summary_table <- StopS %>%
  group_by(AgeRange, Legislation) %>%
  summarise(count = n(),.groups = "drop") %>%
  ungroup()

colnames(summary_table) <- c("AgeRange", "Legislation", "Count")

ggplot(data = summary_table, aes(x = Legislation, y = AgeRange, fill = Count)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(title = "Distribution of age for 401 people stopped by legislation", x = "Types of Legislation", y = "Age range")

```
This heatmap visualizes the distribution of 401 individuals stopped under various legislations, highlighting the prevalence of stops due to "DrugMisuse", especially among those aged 18-24, with over 80 individuals. In contrast, stops under "JusticeFirearms" are comparatively rare, all below 20. The heatmap also indicates a small number of individuals did not disclose their age range, categorized under "Drug Misuse" and "Evidence". Darker colors on the heatmap represent higher numbers of stops, providing a clear visual representation of the data.(79words)

```{r ageLegislationTab, echo=FALSE}
legisType <-  unique(StopS$Legislation)
ageRange <-  unique(StopS$AgeRange)
gender <- unique(StopS$Gender)

legislation_table <- data.frame( 
  Column1 = c(1,2,3,4,5,6,7,8,9),
  Column2 = c(1,2,3,4,5,6,7,8,9),
  Column3 = c(1,2,3,4,5,6,7,8,9),
  Column4 = c(1,2,3,4,5,6,7,8,9),
  Column5 = c(1,2,3,4,5,6,7,8,9))
colnames(legislation_table) <- c(levels(legisType),"All",'proportion')
rownames(legislation_table) <- c('male','female','All','arrest',levels(ageRange))

```

```{r countFrequencies, echo=FALSE}
#count frequencies for each row and column

#count male
select_male_Justi <- subset(StopS, StopS$Legislation == "JusticeFirearms" & Gender == "Male")
male_count_Justi <- nrow(select_male_Justi)
legislation_table[1,1] <- male_count_Justi

select_male_Drug <- subset(StopS, StopS$Legislation == "DrugsMisuse" & Gender == "Male")
male_count_Drug <- nrow(select_male_Drug)
legislation_table[1,2] <- male_count_Drug

select_male_Evi <- subset(StopS, StopS$Legislation == "Evidence" & Gender == "Male")
male_count_Evi <- nrow(select_male_Evi)
legislation_table[1,3] <- male_count_Evi
legislation_table[1,4] <- sum(male_count_Justi,male_count_Drug,male_count_Evi)


#count female
select_female_Justi <- subset(StopS, StopS$Legislation == "JusticeFirearms" & Gender == "Female")
female_count_Justi <- nrow(select_female_Justi)
legislation_table[2,1] <- female_count_Justi

select_female_Drug <- subset(StopS, StopS$Legislation == "DrugsMisuse" & Gender == "Female")
female_count_Drug <- nrow(select_female_Drug)
legislation_table[2,2] <- female_count_Drug

select_female_Evi <- subset(StopS, StopS$Legislation == "Evidence" & Gender == "Female")
female_count_Evi <- nrow(select_female_Evi)
legislation_table[2,3] <- female_count_Evi
legislation_table[2,4] <- sum(female_count_Justi,female_count_Drug,female_count_Evi)


#count arrest

select_arrest_Justi <- subset(StopS, StopS$Legislation == "JusticeFirearms" & Outcome == "Arrest")
arrest_count_Justi <- nrow(select_arrest_Justi)
legislation_table[4,1] <- arrest_count_Justi

select_arrest_Drug <- subset(StopS, StopS$Legislation == "DrugsMisuse" & Outcome == "Arrest")
arrest_count_Drug <- nrow(select_arrest_Drug)
legislation_table[4,2] <- female_count_Drug

select_arrest_Evi <- subset(StopS, StopS$Legislation == "Evidence" & Outcome == "Arrest")
arrest_count_Evi <- nrow(select_arrest_Evi)
legislation_table[4,3] <- arrest_count_Evi
legislation_table[4,4] <- sum(arrest_count_Justi,arrest_count_Drug,arrest_count_Evi)

#count 1-7

select_1t7_Justi <- subset(StopS, StopS$Legislation == "JusticeFirearms" & AgeRange == "1-7")
t7_count_Justi <- nrow(select_1t7_Justi)
legislation_table[5,1] <- t7_count_Justi

select_1t7_Drug <- subset(StopS, StopS$Legislation == "DrugsMisuse" & AgeRange == "1-7")
t7_count_Drug <- nrow(select_1t7_Drug)
legislation_table[5,2] <- t7_count_Drug

select_1t7_Evi <- subset(StopS, StopS$Legislation == "Evidence" & AgeRange == "1-7")
t7_count_Evi <- nrow(select_1t7_Evi)
legislation_table[5,3] <- t7_count_Evi
legislation_table[5,4] <- sum(t7_count_Justi,t7_count_Drug,t7_count_Evi)


#count 10-17

select_10t17_Justi <- subset(StopS, StopS$Legislation == "JusticeFirearms" & AgeRange == "10-17")
t17_count_Justi <- nrow(select_10t17_Justi)
legislation_table[6,1] <- t17_count_Justi

select_10t17_Drug <- subset(StopS, StopS$Legislation == "DrugsMisuse" & AgeRange == "10-17")
t17_count_Drug <- nrow(select_10t17_Drug)
legislation_table[6,2] <- t17_count_Drug

select_10t17_Evi <- subset(StopS, StopS$Legislation == "Evidence" & AgeRange == "10-17")
t17_count_Evi <- nrow(select_10t17_Evi)
legislation_table[6,3] <- t17_count_Evi
legislation_table[6,4] <- sum(t17_count_Justi,t17_count_Drug,t17_count_Evi)


#count 18-24

select_18t24_Justi <- subset(StopS, StopS$Legislation == "JusticeFirearms" & AgeRange == "18-24")
t24_count_Justi <- nrow(select_18t24_Justi)
legislation_table[7,1] <- t24_count_Justi

select_18t24_Drug <- subset(StopS, StopS$Legislation == "DrugsMisuse" & AgeRange == "18-24")
t24_count_Drug <- nrow(select_18t24_Drug)
legislation_table[7,2] <- t24_count_Drug

select_18t24_Evi <- subset(StopS, StopS$Legislation == "Evidence" & AgeRange == "18-24")
t24_count_Evi <- nrow(select_18t24_Evi)
legislation_table[7,3] <- t24_count_Evi
legislation_table[7,4] <- sum(t24_count_Justi,t24_count_Drug,t24_count_Evi)


#count 25-34

select_25t34_Justi <- subset(StopS, StopS$Legislation == "JusticeFirearms" & AgeRange == "25-34")
t34_count_Justi <- nrow(select_25t34_Justi)
legislation_table[8,1] <- t34_count_Justi

select_25t34_Drug <- subset(StopS, StopS$Legislation == "DrugsMisuse" & AgeRange == "25-34")
t34_count_Drug <- nrow(select_25t34_Drug)
legislation_table[8,2] <- t34_count_Drug

select_25t34_Evi <- subset(StopS, StopS$Legislation == "Evidence" & AgeRange == "25-34")
t34_count_Evi <- nrow(select_25t34_Evi)
legislation_table[8,3] <- t34_count_Evi
legislation_table[8,4] <- sum(t34_count_Justi,t34_count_Drug,t34_count_Evi)


#count over 34

select_o34_Justi <- subset(StopS, StopS$Legislation == "JusticeFirearms" & AgeRange == "over 34")
o34_count_Justi <- nrow(select_o34_Justi)
legislation_table[9,1] <- o34_count_Justi

select_o34_Drug <- subset(StopS, StopS$Legislation == "DrugsMisuse" & AgeRange == "over 34")
o34_count_Drug <- nrow(select_o34_Drug)
legislation_table[9,2] <- o34_count_Drug

select_o34_Evi <- subset(StopS, StopS$Legislation == "Evidence" & AgeRange == "over 34")
o34_count_Evi <- nrow(select_o34_Evi)
legislation_table[9,3] <- o34_count_Evi
legislation_table[9,4] <- sum(o34_count_Justi,o34_count_Drug,o34_count_Evi)

prop_male <- round(legislation_table[1,4]/401,2)
legislation_table[1,5] <- prop_male

prop_female <- round(legislation_table[2,4]/401,2)
legislation_table[2,5] <- prop_female

legislation_table[3,5] <- round(legislation_table[3,4]/401,2)

prop_arrest <- round(legislation_table[4,4]/401,2)
legislation_table[4,5] <- prop_arrest

prop_1t7 <- round(legislation_table[5,4]/401,2)
legislation_table[5,5] <- prop_1t7

prop_10t17 <- round(legislation_table[6,4]/401,2)
legislation_table[6,5] <- prop_10t17

prop_18t24 <- round(legislation_table[7,4]/401,2)
legislation_table[7,5] <- prop_18t24

prop_25t34 <- round(legislation_table[8,4]/401,2)
legislation_table[8,5] <- prop_25t34

prop_o34 <- round(legislation_table[9,4]/401,2)
legislation_table[9,5] <- prop_o34


#count all
legislation_table[3,1] <- legislation_table[1,1]+legislation_table[2,1]
legislation_table[3,2] <- legislation_table[1,2]+legislation_table[2,2]
legislation_table[3,3] <- legislation_table[1,3]+legislation_table[2,3]
legislation_table[3,4] <- legislation_table[1,4]+legislation_table[2,4]
legislation_table[3,5] <- legislation_table[1,5]+legislation_table[2,5]

```

```{r creatTab, echo=FALSE}
knitr::kable(legislation_table, format = 'html',caption = "Legislation stops, categorized by gender and age.",align = "c")
cat("\nThis table shows the number of individuals stopped in each group, predominantly males at 355, making up 89% of the total. Only 20% are under arrest. A unique case involves a child aged 1-7 stopped for evidence. The 18-24 age range sees the most stops at 133, around one-third of the total, with 83 due to drug misuse. Determining crime susceptibility across age ranges is challenging, especially for those over 34, due to the broad age range and lack of sufficient data.(82words)")
```


```{r ageEthnicity, echo=FALSE}
Ethinicity <- levels(StopS$OfficerDefEthnic)
ethinic_tab <- data.frame(Column1 = c(1,2,3,4,5,6,7,8,9,0,1),
  Column2 = c(1,2,3,4,5,6,7,8,9,0,1),
  Column3 = c(1,2,3,4,5,6,7,8,9,0,1),
  Column4 = c(1,2,3,4,5,6,7,8,9,0,1),
  Column5 = c(1,2,3,4,5,6,7,8,9,0,1)
)
colnames(ethinic_tab) <- Ethinicity
rownames(ethinic_tab) <- c('male','female','arrest',levels(ageRange),"min AgeRange","All","Proportion")
```


```{r fillDataInEthiTab, echo=FALSE}
#Fill data into Ethinitic table

#count male
select_male_As <- subset(StopS, StopS$OfficerDefEthnic == "Asian" & Gender == "Male")
male_count_As <- nrow(select_male_As)
ethinic_tab[1,1] <- male_count_As

select_male_Bk <- subset(StopS, StopS$OfficerDefEthnic == "Black" & Gender == "Male")
male_count_Bk <- nrow(select_male_Bk)
ethinic_tab[1,2] <- male_count_Bk

select_male_Mx <- subset(StopS, StopS$OfficerDefEthnic == "Mixed" & Gender == "Male")
male_count_Mx <- nrow(select_male_Mx)
ethinic_tab[1,3] <- male_count_Mx

select_male_Ot <- subset(StopS, StopS$OfficerDefEthnic == "Other" & Gender == "Male")
male_count_Ot <- nrow(select_male_Ot)
ethinic_tab[1,4] <- male_count_Ot

select_male_Wt <- subset(StopS, StopS$OfficerDefEthnic == "White" & Gender == "Male")
male_count_Wt <- nrow(select_male_Wt)
ethinic_tab[1,5] <- male_count_Wt


#count female

select_female_As <- subset(StopS, StopS$OfficerDefEthnic == "Asian" & Gender == "Female")
female_count_As <- nrow(select_female_As)
ethinic_tab[2,1] <- female_count_As

select_female_Bk <- subset(StopS, StopS$OfficerDefEthnic == "Black" & Gender == "Female")
female_count_Bk <- nrow(select_female_Bk)
ethinic_tab[2,2] <- female_count_Bk

select_female_Mx <- subset(StopS, StopS$OfficerDefEthnic == "Mixed" & Gender == "Female")
female_count_Mx <- nrow(select_female_Mx)
ethinic_tab[2,3] <- female_count_Mx

select_female_Ot <- subset(StopS, StopS$OfficerDefEthnic == "Other" & Gender == "Female")
female_count_Ot <- nrow(select_female_Ot)
ethinic_tab[2,4] <- female_count_Ot

select_female_Wt <- subset(StopS, StopS$OfficerDefEthnic == "White" & Gender == "Female")
female_count_Wt <- nrow(select_female_Wt)
ethinic_tab[2,5] <- female_count_Wt


#count arrest

select_arrest_As <- subset(StopS, StopS$OfficerDefEthnic == "Asian" & Outcome == "Arrest")
arrest_count_As <- nrow(select_arrest_As)
ethinic_tab[3,1] <- arrest_count_As

select_arrest_Bk <- subset(StopS, StopS$OfficerDefEthnic == "Black" & Outcome == "Arrest")
arrest_count_Bk <- nrow(select_arrest_Bk)
ethinic_tab[3,2] <- arrest_count_Bk

select_arrest_Mx <- subset(StopS, StopS$OfficerDefEthnic == "Mixed" & Outcome == "Arrest")
arrest_count_Mx <- nrow(select_arrest_Mx)
ethinic_tab[3,3] <- arrest_count_Mx

select_arrest_Ot <- subset(StopS, StopS$OfficerDefEthnic == "Other" & Outcome == "Arrest")
arrest_count_Ot <- nrow(select_arrest_Ot)
ethinic_tab[3,4] <- arrest_count_Ot

select_arrest_Wt <- subset(StopS, StopS$OfficerDefEthnic == "White" & Outcome == "Arrest")
arrest_count_Wt <- nrow(select_arrest_Wt)
ethinic_tab[3,5] <- arrest_count_Wt


#count 1-7

select_1t7_As <- subset(StopS, StopS$OfficerDefEthnic == "Asian" & AgeRange == "1-7")
t7_count_As <- nrow(select_1t7_As)
ethinic_tab[4,1] <- t7_count_As

select_1t7_Bk <- subset(StopS, StopS$OfficerDefEthnic == "Black" & AgeRange == "1-7")
t7_count_Bk <- nrow(select_1t7_Bk)
ethinic_tab[4,2] <- t7_count_Bk

select_1t7_Mx <- subset(StopS, StopS$OfficerDefEthnic == "Mixed" & AgeRange == "1-7")
t7_count_Mx <- nrow(select_1t7_Mx)
ethinic_tab[4,3] <- t7_count_Mx

select_1t7_Ot <- subset(StopS, StopS$OfficerDefEthnic == "Other" & AgeRange == "1-7")
t7_count_Ot <- nrow(select_1t7_Ot)
ethinic_tab[4,4] <- t7_count_Ot

select_1t7_Wt <- subset(StopS, StopS$OfficerDefEthnic == "White" & AgeRange == "1-7")
t7_count_Wt <- nrow(select_1t7_Wt)
ethinic_tab[4,5] <- t7_count_Wt


#count 10-17

select_10t17_As <- subset(StopS, StopS$OfficerDefEthnic == "Asian" & AgeRange == "10-17")
t17_count_As <- nrow(select_10t17_As)
ethinic_tab[5,1] <- t17_count_As

select_10t17_Bk <- subset(StopS, StopS$OfficerDefEthnic == "Black" & AgeRange == "10-17")
t17_count_Bk <- nrow(select_10t17_Bk)
ethinic_tab[5,2] <- t17_count_Bk

select_10t17_Mx <- subset(StopS, StopS$OfficerDefEthnic == "Mixed" & AgeRange == "10-17")
t17_count_Mx <- nrow(select_10t17_Mx)
ethinic_tab[5,3] <- t17_count_Mx

select_10t17_Ot <- subset(StopS, StopS$OfficerDefEthnic == "Other" & AgeRange == "10-17")
t17_count_Ot <- nrow(select_10t17_Ot)
ethinic_tab[5,4] <- t17_count_Ot

select_10t17_Wt <- subset(StopS, StopS$OfficerDefEthnic == "White" & AgeRange == "10-17")
t17_count_Wt <- nrow(select_10t17_Wt)
ethinic_tab[5,5] <- t17_count_Wt


#count 18-24

select_18t24_As <- subset(StopS, StopS$OfficerDefEthnic == "Asian" & AgeRange == "18-24")
t24_count_As <- nrow(select_18t24_As)
ethinic_tab[6,1] <- t24_count_As

select_18t24_Bk <- subset(StopS, StopS$OfficerDefEthnic == "Black" & AgeRange == "18-24")
t24_count_Bk <- nrow(select_18t24_Bk)
ethinic_tab[6,2] <- t24_count_Bk

select_18t24_Mx <- subset(StopS, StopS$OfficerDefEthnic == "Mixed" & AgeRange == "18-24")
t24_count_Mx <- nrow(select_18t24_Mx)
ethinic_tab[6,3] <- t24_count_Mx

select_18t24_Ot <- subset(StopS, StopS$OfficerDefEthnic == "Other" & AgeRange == "18-24")
t24_count_Ot <- nrow(select_18t24_Ot)
ethinic_tab[6,4] <- t24_count_Ot

select_18t24_Wt <- subset(StopS, StopS$OfficerDefEthnic == "White" & AgeRange == "18-24")
t24_count_Wt <- nrow(select_18t24_Wt)
ethinic_tab[6,5] <- t24_count_Wt


#count 25-34

select_25t34_As <- subset(StopS, StopS$OfficerDefEthnic == "Asian" & AgeRange == "25-34")
t34_count_As <- nrow(select_25t34_As)
ethinic_tab[7,1] <- t34_count_As

select_25t34_Bk <- subset(StopS, StopS$OfficerDefEthnic == "Black" & AgeRange == "25-34")
t34_count_Bk <- nrow(select_25t34_Bk)
ethinic_tab[7,2] <- t34_count_Bk

select_25t34_Mx <- subset(StopS, StopS$OfficerDefEthnic == "Mixed" & AgeRange == "25-34")
t34_count_Mx <- nrow(select_25t34_Mx)
ethinic_tab[7,3] <- t34_count_Mx

select_25t34_Ot <- subset(StopS, StopS$OfficerDefEthnic == "Other" & AgeRange == "25-34")
t34_count_Ot <- nrow(select_25t34_Ot)
ethinic_tab[7,4] <- t34_count_Ot

select_25t34_Wt <- subset(StopS, StopS$OfficerDefEthnic == "White" & AgeRange == "25-34")
t34_count_Wt <- nrow(select_25t34_Wt)
ethinic_tab[7,5] <- t34_count_Wt


#count over 34

select_o34_As <- subset(StopS, StopS$OfficerDefEthnic == "Asian" & AgeRange == "over 34")
o34_count_As <- nrow(select_o34_As)
ethinic_tab[8,1] <- o34_count_As

select_o34_Bk <- subset(StopS, StopS$OfficerDefEthnic == "Black" & AgeRange == "over 34")
o34_count_Bk <- nrow(select_o34_Bk)
ethinic_tab[8,2] <- o34_count_Bk

select_o34_Mx <- subset(StopS, StopS$OfficerDefEthnic == "Mixed" & AgeRange == "over 34")
o34_count_Mx <- nrow(select_o34_Mx)
ethinic_tab[8,3] <- o34_count_Mx

select_o34_Ot <- subset(StopS, StopS$OfficerDefEthnic == "Other" & AgeRange == "over 34")
o34_count_Ot <- nrow(select_o34_Ot)
ethinic_tab[8,4] <- o34_count_Ot

select_o34_Wt <- subset(StopS, StopS$OfficerDefEthnic == "White" & AgeRange == "over 34")
o34_count_Wt <- nrow(select_o34_Wt)
ethinic_tab[8,5] <- o34_count_Wt


#count all
all_as <- male_count_As + female_count_As
all_bk <- male_count_Bk + female_count_Bk
all_mx <- male_count_Mx + female_count_Mx
all_ot <- male_count_Ot + female_count_Ot
all_wt <- male_count_Wt + female_count_Wt


#count proportion
prop_as <- round(all_as/401,2)
prop_bk <- round(all_bk/401,2)
prop_mx <- round(all_mx/401,2)
prop_ot <- round(all_ot/401,2)
prop_wt <- round(all_wt/401,2)
```

```{r selectMinAge, echo=FALSE}
#min AgeRange

ethinic_tab[9,1] <- "10-17"

ethinic_tab[9,2] <- "10-17"

ethinic_tab[9,3] <- "10-17"

ethinic_tab[9,4] <- "18-24"

ethinic_tab[9,5] <- "1-7"

ethinic_tab[10,1] <- all_as
ethinic_tab[10,2] <- all_bk
ethinic_tab[10,3] <- all_mx
ethinic_tab[10,4] <- all_ot
ethinic_tab[10,5] <- all_wt

ethinic_tab[11,1] <- prop_as
ethinic_tab[11,2] <- prop_bk
ethinic_tab[11,3] <- prop_mx
ethinic_tab[11,4] <- prop_ot
ethinic_tab[11,5] <- prop_wt


```


```{r createEthicTab,echo=FALSE}
knitr::kable(ethinic_tab,caption = "Stopped People Ethinicity, by AgeRange,Gender",align = "c")
cat("\nThis table presents the ethnicity of individuals stopped, with White being the majority at 70% (282 people), likely due to data collection in the UK. Black and Asian follow at 13% and 11%, respectively. The minimum AgeRange for White is 1-7, the lowest overall, while for Asian, Black, and Mixed, it is 10-17. The 'Other' category starts from 18-24, but the data may not accurately reflect the real situation due to its small size of only four entries. (78words)")
```



```{r summaryTable, echo=FALSE}
library("kableExtra")
summary(StopS)

summary_table <- data.frame(
  Male = 1:26,
  Female = 1:26,
  All = 1:26
)
colnames(summary_table) <- c("Male","Female","All")
levels(StopS$Type)
levels(StopS$AgeRange)
levels(StopS$OfficerDefEthnic)
levels(StopS$Legislation)
outcome_type <- unique(StopS$Outcome)
levels(outcome_type)
```



