---
title: "AnalysisData"
author: '2140365'
date: "2023-10-28"
output: html_document
---

```{r GDP-povertyDt, echo=FALSE,include=FALSE}
library(rio)
library(dplyr)
GDP_poverty <- read.csv("SDG1_2023/GDP-povertyPercentages.csv")
GDP_poverty <- GDP_poverty%>%mutate(
  country = factor(country),
  code = factor(code),
  year = factor(year)
)
GDP_poverty_5 <- filter(GDP_poverty, country %in% c(
  "Germany","Egypt, Arab Rep.","Nigeria","Australia","United Kingdom"
)) 
```
'GDP_poverty' listed the the percentage_poverty of the country, and its GDP in given year, i dont really understand what 'pct_poverty' refers to. We can compare 5 countries' GDP trend in years and the relationship with 'pct_poverty'


```{r poverty-surveysdt ,echo=FALSE,include=FALSE}
poverty_survey <- read.csv("SDG1_2023/Poverty-surveys-per-decade.csv")
poverty_survey <- poverty_survey%>%mutate(
  Entity = factor(Entity),
  Code = factor(Code),
  Year = factor(Year)
)
poverty_survey_5 <- filter(poverty_survey, Entity %in% c(
  "Germany","Egypt","Nigeria","Australia","United Kingdom"
)) 
```
'poverty_survey' mainly listed the number of. surveys have done, maybe can explain why some of them have survey and some of them not?


```{r share-extreme-poverty,echo=FALSE,include=FALSE}
share_poverty <- read.csv("SDG1_2023/share-extreme-poverty3.65dollarDay.csv")
share_poverty <- share_poverty%>%mutate(
  Entity = factor(Entity),
  Code = factor(Code),
  Year = factor(Year)
)
share_poverty_5 <- filter(share_poverty,Entity%in% c(
  "Germany","Egypt","Nigeria","Australia","United Kingdom"
))
```
'pctBelowPovertyLine' can be discussed in this dataset, talk about its trend with years


```{r countryStuntingData,echo=FALSE,include=FALSE}
load("SDG2/StuntingCountryOview.Rda")
```
record of the detailds of Stunting data, identified by Code,


```{r metaData-country,echo=FALSE,include=FALSE}
metadata_country <- read.csv("SDG2/Metadata-Country-Stunting.csv")
metadata_country <- metadata_country%>%mutate(
  CountryCode = factor(CountryCode),
  CountryName = factor(CountryName),
  Region = factor(Region),
  IncomeGroup = factor(IncomeGroup)
)
metadate_country_5 <- filter(metadata_country, CountryName%in% c(
  "Fiji","Egypt, Arab Rep.","Nigeria","Australia","United Kingdom"
))
```
we should mainly read the Special notes, find the relation between regions and IncomeGroup?


```{r metadata-underfed,echo=FALSE,include=FALSE}
country_underFed<- read.csv("SDG2/Metadata-Country-UnderFed.csv")
country_underFed <- country_underFed%>%mutate(
  CountryCode = factor(CountryCode),
  CountryName = factor(CountryName),
  Region = factor(Region),
  IncomeGroup = factor(IncomeGroup)
)
country_underFed_5 <- filter(country_underFed, CountryName%in% c(
  "Fiji","Egypt, Arab Rep.","Nigeria","Australia","United Kingdom"
))
```
I dont understand this actually


```{r metadata-region,echo=FALSE,include=FALSE}
metadata_region <- read.csv("SDG2/Metadata-Regions-Stunting.csv")
colnames(metadata_region) <- c("Code","IncomeGroup","Region","SpecialNotes")
metadata_region <- metadata_region%>%mutate(
  Code = factor(Code),
  IncomeGroup = factor(IncomeGroup),
  Region = factor(Region)
)
metadata_region_5 <- filter(metadata_country, CountryName%in% c(
  "Fiji","Egypt, Arab Rep.","Nigeria","Australia","United Kingdom"
))
```
just read the special notes i think


```{r StuntingU5, echo=FALSE, include=FALSE}
StuntingU5yrs <- read.csv("SDG2/StuntingU5yrs2023.csv")
StuntingU5yrs <- StuntingU5yrs%>%mutate(
  Country = factor(Country),
  Code = factor(Code),
  year = factor(year)
)
StuntingU5yrs_5 <- filter(StuntingU5yrs, Country%in% c(
  "Fiji","Egypt, Arab Rep.","Nigeria","Australia","United Kingdom"
))
```
Talk about the rate of StuntingU5yrs wrt years, but idk why there is no UK


```{r UnderFed, echo=FALSE, include=FALSE}
Underfed <- read.csv("SDG2/Underfed.csv")
Underfed_5 <- filter(Underfed, Entity%in%c("Fiji","Egypt","Nigeria","Australia","United Kingdom of Great Britain and Northern Ireland"))
```
Underfed.csv considered whole continent as a entity. we may need to consider the comparison between countries with Continent. 




